Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-11 podc16/paper2.tex
work on graph linearization

The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$). For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.

Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$. Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order. For example, if $V\left( g
\right)=\aset{c_1,c_2,c_3,c_4}$
and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.

A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
Let $\opn{TOs}\left( g \right)$ be the set of total orders $\leq$ such that $\preceq_g \subseteq \leq$

Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.

When $\hat{o}$ is a total order in $\opn{TOs}\left( \hat{g} \right)$, define the total order $\leq_{\hat{o}}^g$ on $V\left( g \right)$ such that $v_1 \leq_{\hat{o}}^g v_2$ iff (1) there exists $V_1,V_2\in V\left(\hat{g}\right)$ such that $v_1\in V_1$, $v_2\in V_2$, and $\left( V_1,V_2 \right)\in \hat{o}$, or (2) there exists $V\in V\left(\hat{g}\right)$ such that $v_1,v_2\in V\left( V \right)$ and $\left( v_1,v_2 \right)\in f\left[ V\left( V \right) \right]$.
We say that $\leq_{\hat{o}}^g$ is a \emph{linearization} of $g$.

Property: if $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, $E\left( g_1 \right)\subseteq E\left( g_2 \right)$, $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$, then $\bot\bullet l1 \preceq \bot\bullet l2$.

>>>>

\textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$). For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.

\textbf{Graph of strongly connected components.} Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ (represented as sets of vertices) and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$. Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order. For example, if $V\left( g
\right)=\aset{c_1,c_2,c_3,c_4}$
and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.

\textbf{Orders induced by a graph.} A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.

\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.

\textbf{Linearizations of a dependency graph.}
When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, define the total order $l\left(\hat{o},g\right)$ on $V\left( g \right)$ such that $\left(v_1,v_2\right) \in l\left(\hat{o},g\right)$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
$v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
We say that $l\left(\hat{o},g\right)$ is a \emph{linearization} of $g$.

\textbf{Lemma}: 
Assume that: $g_1$ and $g_2$ are two dependency graphs; 
for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
$V\left( \hat{g_1} \right)\subseteq V\left( \hat{g_2} \right)$; 
$E\left( \hat{g_1} \right)\subseteq E\left( \hat{g_2} \right)$; 
$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
Then, $\bot\bullet l1 \preceq \bot\bullet l2$.
